Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks

### 2023-01-06 paper-1/manuscript.tex
Rev: Tighten explanation of mean R for D

A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity. Here, we used the average trial response $R_j^S$ of neuron $j$ to a given stimulus $S$, in number of spikes, as a measure of the presynaptic drive. Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at $V_m = V_{rest}$ (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$. Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $R_j^S k x_j^S$. For this estimate, we ignored added depression due to repeated spikes within a trial, instead defining $x_j^S$ as the mean depression at the start of trials. Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in a given neuron $i$ as
\begin{equation}
    D_i^S = \sum_{j \in pre} R_j^S k (1-x_j^S)
\end{equation}
Notice that $D$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $V_{TA}$, albeit more approximate. Lastly, since we contrasted this measure between conditions (see below), we aimed to avoid effects purely due to differences in presynaptic activity. Therefore, we calculated $R^S$ as the average response across $S$ trials in all conditions, leaving the contrast entirely in the hands of $x^S$.

>>>>

A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity. In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $R_j^S$ of a neuron $j$ to a given stimulus $S$ across all sequences as a measure of the presynaptic drive. Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at $V_m = V_{rest}$ (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$. Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $R_j^S k x_j^S$, where we ignored added depression due to repeated spikes within a trial and instead defined $x_j^S$ as the mean depression at the start of trials. Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in a given neuron $i$ as
\begin{equation}
    D_i^S = \sum_{j \in pre} R_j^S k (1-x_j^S)
\end{equation}
Notice that $D$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $V_{TA}$, albeit more approximate.
